---
output:
  html_document: default
  pdf_document: default
  code_folding: "hide"
---
```{r include = FALSE}
knitr::opts_chunk$set(echo=FALSE, results="hide")
setwd("D:/ECON613/A1")
rm(list = ls())
library(dplyr)
library(tidyr)
library(reshape2)
library(data.table)
library(xtable)
datstu=read.csv('datstu.csv')
datsss=read.csv('datsss.csv')
datjss=read.csv('datjss.csv')
```

# Exercise 1

```{r echo=TRUE, results="show"}
#Exercise 1 Missing Data----
##Number of Students----
length(datstu$X)
##Number of Schools----
length(na.omit(unique(datsss$schoolcode)))
length(na.omit(unique(unlist(datstu[5:10]))))
##Number of Programs----
length(unique(na.omit(unlist(datstu[11:16]))))
##Number of Choices----
datstu=datstu %>% mutate(choice1=paste(schoolcode1, choicepgm1, sep=" "),
                         choice2=paste(schoolcode2, choicepgm2, sep=" "),
                         choice3=paste(schoolcode3, choicepgm3, sep=" "),
                         choice4=paste(schoolcode4, choicepgm4, sep=" "),
                         choice5=paste(schoolcode5, choicepgm5, sep=" "),
                         choice6=paste(schoolcode6, choicepgm6, sep=" "))
dat=select(datstu, choice1, choice2, choice3, choice4, choice5, choice6)
dat_long=gather(dat, 'key', 'value')
dat_long <- subset(dat_long, value!= "NA ")
length(unique(na.omit(dat_long$value)))
##Number of Missing Test Scores----
sum(is.na(datstu$score))
##Number of Same School, Different Programs----
for (i in 1:nrow(datstu)){
datstu$sameschool[i]=length(unique(na.omit(unlist(datstu[i,5:10]))))
}
sum(datstu$sameschool<6-rowSums(is.na(datstu[5:10])))
#sum=result$`combine$dat2.program`
##Apply to less than 6 choices----
length(which(rowSums(dat=="NA ")!=0))

```
| Number of students: 340823
| Number of schools(Junior High): 898
| Number of schools(Senior HIgh that students apply): 640
| Number of programs: 33
| Number of choices (school,program): 3085
| Missing test score: 179887
| Apply to the same school (different programs): 120071
| Apply to less than 6 choices: 17720
| #Note here some of the results may be different due to different manipulations on NA values in original dataset 


# Exercise 2

```{r echo=TRUE, results="show"}
#Exercise 2 Data----
datstu$admitted_schoolcode=ifelse(datstu$rankplace==1, datstu$schoolcode1,
                       ifelse(datstu$rankplace==2, datstu$schoolcode2,
                              ifelse(datstu$rankplace==3, datstu$schoolcode3,
                                     ifelse(datstu$rankplace==4, datstu$schoolcode4,
                                            ifelse(datstu$rankplace==5, datstu$schoolcode5,
                                                   ifelse(datstu$rankplace==6, datstu$schoolcode6, NA))))))
datstu$admitted=ifelse(datstu$rankplace==1, datstu$choice1,
                       ifelse(datstu$rankplace==2, datstu$choice2,
                              ifelse(datstu$rankplace==3, datstu$choice3,
                                     ifelse(datstu$rankplace==4, datstu$choice4,
                                            ifelse(datstu$rankplace==5, datstu$choice5,
                                                   ifelse(datstu$rankplace==6, datstu$choice6, NA))))))
dataset=datstu %>% group_by(admitted) %>% summarise(schoolcode=admitted_schoolcode, minscore=min(score), average=mean(score), number=n(), .groups = 'drop')
dataset=dataset %>% 
  rename(
    school_program = admitted
  )
dataset=unique(dataset)
datsss$X=NULL
datsss=unique(datsss)
datafinal<-merge(x=dataset,y=datsss,by="schoolcode",all.x=TRUE)
datafinal=na.omit(datafinal)
#Another way to left join
datatest=left_join(dataset, datsss, by="schoolcode")
datatest=na.omit(datatest)
datatest_final=select(datatest, school_program, sssdistrict, ssslat, ssslong, minscore, average, number)
datatest_final=datatest_final %>% 
  rename(
    district=sssdistrict,
    latitude=ssslat,
    longitude=ssslong,
    cutoff=minscore,
    quality=average,
    size=number
  )
head(datatest_final, 20)
```

# Exercise 3

```{r echo=TRUE, results="show"}
#Exercise 3 Distance----
datjss=select(datjss, -X)
datstujss=left_join(datstu, datjss, by="jssdistrict")
dat_school=select(datstu, X, choice1:choice6)
dat_school=gather(dat_school, 'key', 'value', -X)
dat_school=left_join(dat_school, datstujss, by=c("X"="X"))
dat_school=select(dat_school, X, key, value, jssdistrict, point_x, point_y)
dat_school=cbind(dat_school, colsplit(dat_school$value," ",c("schoolcode","program")))

m=dat_school%>%group_by(X)

datstujsssss=left_join(dat_school, datsss, by=c("schoolcode"="schoolcode"))
datstujsssss$distance=sqrt(
  (69.172*(datstujsssss$ssslong-datstujsssss$point_x)*cos(datstujsssss$point_y/57.3))^2+(69.172*(datstujsssss$ssslat-datstujsssss$point_y)^2)
)
datstujsssss_omit=datstujsssss[!is.na(datstujsssss$distance),]

datstujsssss_omit=select(datstujsssss_omit, jssdistrict, point_x, point_y, schoolname, sssdistrict, ssslat, ssslong, distance)
datstujsssss_omit=datstujsssss_omit %>% 
  rename(
    jsslat=point_y,
    jsslong=point_x,
    sssname=schoolname,
  )

head(datstujsssss_omit, 20)
```

# Exercise 4
```{r echo=TRUE, results="show"}
#Exercise 4 Descriptive Characteristics----
datstujsssss=left_join(datstujsssss, dataset, by=c("value"="school_program"))
score=data.frame(datstu[,1:2])
datstujsssss=left_join(datstujsssss, score, by=c("X"="X"))


#choice1
rank1=na.omit(datstujsssss[datstujsssss$key=="choice1",])
sd_cutoff_choice1=sd(rank1$minscore)
mean_cutoff_choice1=mean(rank1$minscore)
sd_quality_choice1=sd(rank1$average)
mean_quality_choice1=mean(rank1$average)
sd_distance_choice1=sd(rank1$distance)
mean_distance_choice1=mean(rank1$distance)

#choice2
rank2=na.omit(datstujsssss[datstujsssss$key=="choice2",])
sd_cutoff_choice2=sd(rank2$minscore)
mean_cutoff_choice2=mean(rank2$minscore)
sd_quality_choice2=sd(rank2$average)
mean_quality_choice2=mean(rank2$average)
sd_distance_choice2=sd(rank2$distance)
mean_distance_choice2=mean(rank2$distance)

#choice3
rank3=na.omit(datstujsssss[datstujsssss$key=="choice3",])
sd_cutoff_choice3=sd(rank3$minscore)
mean_cutoff_choice3=mean(rank3$minscore)
sd_quality_choice3=sd(rank3$average)
mean_quality_choice3=mean(rank3$average)
sd_distance_choice3=sd(rank3$distance)
mean_distance_choice3=mean(rank3$distance)

#choice4
rank4=na.omit(datstujsssss[datstujsssss$key=="choice4",])
sd_cutoff_choice4=sd(rank4$minscore)
mean_cutoff_choice4=mean(rank4$minscore)
sd_quality_choice4=sd(rank4$average)
mean_quality_choice4=mean(rank4$average)
sd_distance_choice4=sd(rank4$distance)
mean_distance_choice4=mean(rank4$distance)

#choice5
rank5=na.omit(datstujsssss[datstujsssss$key=="choice5",])
sd_cutoff_choice5=sd(rank5$minscore)
mean_cutoff_choice5=mean(rank5$minscore)
sd_quality_choice5=sd(rank5$average)
mean_quality_choice5=mean(rank5$average)
sd_distance_choice5=sd(rank5$distance)
mean_distance_choice5=mean(rank5$distance)

#choice6
rank6=na.omit(datstujsssss[datstujsssss$key=="choice6",])
sd_cutoff_choice6=sd(rank6$minscore)
mean_cutoff_choice6=mean(rank6$minscore)
sd_quality_choice6=sd(rank6$average)
mean_quality_choice6=mean(rank6$average)
sd_distance_choice6=sd(rank6$distance)
mean_distance_choice6=mean(rank6$distance)

mean_sd_table=cbind(c(sd_cutoff_choice1, sd_cutoff_choice2, sd_cutoff_choice3, sd_cutoff_choice4, sd_cutoff_choice5, sd_cutoff_choice6), c(mean_cutoff_choice1, mean_cutoff_choice2, mean_cutoff_choice3, mean_cutoff_choice4, mean_cutoff_choice5, mean_cutoff_choice6), c(sd_quality_choice1, sd_quality_choice2, sd_quality_choice3, sd_quality_choice4, sd_quality_choice5, sd_quality_choice6), c(mean_quality_choice1, mean_quality_choice2, mean_quality_choice3, mean_quality_choice4, mean_quality_choice5, mean_quality_choice6), c(sd_distance_choice1, sd_distance_choice2, sd_distance_choice3, sd_distance_choice4, sd_distance_choice5, sd_distance_choice6), c(mean_distance_choice1, mean_distance_choice2, mean_distance_choice3, mean_distance_choice4, mean_distance_choice5, mean_distance_choice6))

colnames(mean_sd_table) = c("sd_cutoff", "mean_cutoff","sd_quality","mean_quality", "sd_distance", "mean_distance")
rownames(mean_sd_table) = c("choice1", "choice2","choice3","choice4", "choice5", "choice6")




#datstu_omit=datstu[!is.na(datstu$score),]

#choice1_quantile
rank1_q <- rank1 %>% 
  mutate(group = cut(score, c(quantile(score)),
                     labels = 1:4, include.lowest=T))
rank1_q1=rank1_q[rank1_q$group==1, ]
rank1_q2=rank1_q[rank1_q$group==2, ]
rank1_q3=rank1_q[rank1_q$group==3, ]
rank1_q4=rank1_q[rank1_q$group==4, ]

sd_cutoff_r1q1=sd(rank1_q1$minscore)
mean_cutoff_r1q1=mean(rank1_q1$minscore)
sd_cutoff_r1q2=sd(rank1_q2$minscore)
mean_cutoff_r1q2=mean(rank1_q2$minscore)
sd_cutoff_r1q3=sd(rank1_q3$minscore)
mean_cutoff_r1q3=mean(rank1_q3$minscore)
sd_cutoff_r1q4=sd(rank1_q4$minscore)
mean_cutoff_r1q4=mean(rank1_q4$minscore)


sd_quality_r1q1=sd(rank1_q1$average)
mean_quality_r1q1=mean(rank1_q1$average)
sd_quality_r1q2=sd(rank1_q2$average)
mean_quality_r1q2=mean(rank1_q2$average)
sd_quality_r1q3=sd(rank1_q3$average)
mean_quality_r1q3=mean(rank1_q3$average)
sd_quality_r1q4=sd(rank1_q4$average)
mean_quality_r1q4=mean(rank1_q4$average)

sd_distance_r1q1=sd(rank1_q1$distance)
mean_distance_r1q1=mean(rank1_q1$distance)
sd_distance_r1q2=sd(rank1_q2$distance)
mean_distance_r1q2=mean(rank1_q2$distance)
sd_distance_r1q3=sd(rank1_q3$distance)
mean_distance_r1q3=mean(rank1_q3$distance)
sd_distance_r1q4=sd(rank1_q4$distance)
mean_distance_r1q4=mean(rank1_q4$distance)



#choice2_quantile
rank2_q <- rank2 %>% 
  mutate(group = cut(score, c(quantile(score)),
                     labels = 1:4, include.lowest=T))
rank2_q1=rank2_q[rank2_q$group==1, ]
rank2_q2=rank2_q[rank2_q$group==2, ]
rank2_q3=rank2_q[rank2_q$group==3, ]
rank2_q4=rank2_q[rank2_q$group==4, ]

sd_cutoff_r2q1=sd(rank2_q1$minscore)
mean_cutoff_r2q1=mean(rank2_q1$minscore)
sd_cutoff_r2q2=sd(rank2_q2$minscore)
mean_cutoff_r2q2=mean(rank2_q2$minscore)
sd_cutoff_r2q3=sd(rank2_q3$minscore)
mean_cutoff_r2q3=mean(rank2_q3$minscore)
sd_cutoff_r2q4=sd(rank2_q4$minscore)
mean_cutoff_r2q4=mean(rank2_q4$minscore)


sd_quality_r2q1=sd(rank2_q1$average)
mean_quality_r2q1=mean(rank2_q1$average)
sd_quality_r2q2=sd(rank2_q2$average)
mean_quality_r2q2=mean(rank2_q2$average)
sd_quality_r2q3=sd(rank2_q3$average)
mean_quality_r2q3=mean(rank2_q3$average)
sd_quality_r2q4=sd(rank2_q4$average)
mean_quality_r2q4=mean(rank2_q4$average)

sd_distance_r2q1=sd(rank2_q1$distance)
mean_distance_r2q1=mean(rank2_q1$distance)
sd_distance_r2q2=sd(rank2_q2$distance)
mean_distance_r2q2=mean(rank2_q2$distance)
sd_distance_r2q3=sd(rank2_q3$distance)
mean_distance_r2q3=mean(rank2_q3$distance)
sd_distance_r2q4=sd(rank2_q4$distance)
mean_distance_r2q4=mean(rank2_q4$distance)

#choice3_quantile
rank3_q <- rank3 %>% 
  mutate(group = cut(score, c(quantile(score)),
                     labels = 1:4, include.lowest=T))
rank3_q1=rank3_q[rank3_q$group==1, ]
rank3_q2=rank3_q[rank3_q$group==2, ]
rank3_q3=rank3_q[rank3_q$group==3, ]
rank3_q4=rank3_q[rank3_q$group==4, ]

sd_cutoff_r3q1=sd(rank3_q1$minscore)
mean_cutoff_r3q1=mean(rank3_q1$minscore)
sd_cutoff_r3q2=sd(rank3_q2$minscore)
mean_cutoff_r3q2=mean(rank3_q2$minscore)
sd_cutoff_r3q3=sd(rank3_q3$minscore)
mean_cutoff_r3q3=mean(rank3_q3$minscore)
sd_cutoff_r3q4=sd(rank3_q4$minscore)
mean_cutoff_r3q4=mean(rank3_q4$minscore)


sd_quality_r3q1=sd(rank3_q1$average)
mean_quality_r3q1=mean(rank3_q1$average)
sd_quality_r3q2=sd(rank3_q2$average)
mean_quality_r3q2=mean(rank3_q2$average)
sd_quality_r3q3=sd(rank3_q3$average)
mean_quality_r3q3=mean(rank3_q3$average)
sd_quality_r3q4=sd(rank3_q4$average)
mean_quality_r3q4=mean(rank3_q4$average)

sd_distance_r3q1=sd(rank3_q1$distance)
mean_distance_r3q1=mean(rank3_q1$distance)
sd_distance_r3q2=sd(rank3_q2$distance)
mean_distance_r3q2=mean(rank3_q2$distance)
sd_distance_r3q3=sd(rank3_q3$distance)
mean_distance_r3q3=mean(rank3_q3$distance)
sd_distance_r3q4=sd(rank3_q4$distance)
mean_distance_r3q4=mean(rank3_q4$distance)

#choice4_quantile
rank4_q <- rank4 %>% 
  mutate(group = cut(score, c(quantile(score)),
                     labels = 1:4, include.lowest=T))
rank4_q1=rank4_q[rank4_q$group==1, ]
rank4_q2=rank4_q[rank4_q$group==2, ]
rank4_q3=rank4_q[rank4_q$group==3, ]
rank4_q4=rank4_q[rank4_q$group==4, ]

sd_cutoff_r4q1=sd(rank4_q1$minscore)
mean_cutoff_r4q1=mean(rank4_q1$minscore)
sd_cutoff_r4q2=sd(rank4_q2$minscore)
mean_cutoff_r4q2=mean(rank4_q2$minscore)
sd_cutoff_r4q3=sd(rank4_q3$minscore)
mean_cutoff_r4q3=mean(rank4_q3$minscore)
sd_cutoff_r4q4=sd(rank4_q4$minscore)
mean_cutoff_r4q4=mean(rank4_q4$minscore)


sd_quality_r4q1=sd(rank4_q1$average)
mean_quality_r4q1=mean(rank4_q1$average)
sd_quality_r4q2=sd(rank4_q2$average)
mean_quality_r4q2=mean(rank4_q2$average)
sd_quality_r4q3=sd(rank4_q3$average)
mean_quality_r4q3=mean(rank4_q3$average)
sd_quality_r4q4=sd(rank4_q4$average)
mean_quality_r4q4=mean(rank4_q4$average)

sd_distance_r4q1=sd(rank4_q1$distance)
mean_distance_r4q1=mean(rank4_q1$distance)
sd_distance_r4q2=sd(rank4_q2$distance)
mean_distance_r4q2=mean(rank4_q2$distance)
sd_distance_r4q3=sd(rank4_q3$distance)
mean_distance_r4q3=mean(rank4_q3$distance)
sd_distance_r4q4=sd(rank4_q4$distance)
mean_distance_r4q4=mean(rank4_q4$distance)

#choice5_quantile
rank5_q <- rank5 %>% 
  mutate(group = cut(score, c(quantile(score)),
                     labels = 1:4, include.lowest=T))
rank5_q1=rank5_q[rank5_q$group==1, ]
rank5_q2=rank5_q[rank5_q$group==2, ]
rank5_q3=rank5_q[rank5_q$group==3, ]
rank5_q4=rank5_q[rank5_q$group==4, ]

sd_cutoff_r5q1=sd(rank5_q1$minscore)
mean_cutoff_r5q1=mean(rank5_q1$minscore)
sd_cutoff_r5q2=sd(rank5_q2$minscore)
mean_cutoff_r5q2=mean(rank5_q2$minscore)
sd_cutoff_r5q3=sd(rank5_q3$minscore)
mean_cutoff_r5q3=mean(rank5_q3$minscore)
sd_cutoff_r5q4=sd(rank5_q4$minscore)
mean_cutoff_r5q4=mean(rank5_q4$minscore)


sd_quality_r5q1=sd(rank5_q1$average)
mean_quality_r5q1=mean(rank5_q1$average)
sd_quality_r5q2=sd(rank5_q2$average)
mean_quality_r5q2=mean(rank5_q2$average)
sd_quality_r5q3=sd(rank5_q3$average)
mean_quality_r5q3=mean(rank5_q3$average)
sd_quality_r5q4=sd(rank5_q4$average)
mean_quality_r5q4=mean(rank5_q4$average)

sd_distance_r5q1=sd(rank5_q1$distance)
mean_distance_r5q1=mean(rank5_q1$distance)
sd_distance_r5q2=sd(rank5_q2$distance)
mean_distance_r5q2=mean(rank5_q2$distance)
sd_distance_r5q3=sd(rank5_q3$distance)
mean_distance_r5q3=mean(rank5_q3$distance)
sd_distance_r5q4=sd(rank5_q4$distance)
mean_distance_r5q4=mean(rank5_q4$distance)

#choice6_quantile
rank6_q <- rank6 %>% 
  mutate(group = cut(score, c(quantile(score)),
                     labels = 1:4, include.lowest=T))
rank6_q1=rank6_q[rank6_q$group==1, ]
rank6_q2=rank6_q[rank6_q$group==2, ]
rank6_q3=rank6_q[rank6_q$group==3, ]
rank6_q4=rank6_q[rank6_q$group==4, ]

sd_cutoff_r6q1=sd(rank6_q1$minscore)
mean_cutoff_r6q1=mean(rank6_q1$minscore)
sd_cutoff_r6q2=sd(rank6_q2$minscore)
mean_cutoff_r6q2=mean(rank6_q2$minscore)
sd_cutoff_r6q3=sd(rank6_q3$minscore)
mean_cutoff_r6q3=mean(rank6_q3$minscore)
sd_cutoff_r6q4=sd(rank6_q4$minscore)
mean_cutoff_r6q4=mean(rank6_q4$minscore)


sd_quality_r6q1=sd(rank6_q1$average)
mean_quality_r6q1=mean(rank6_q1$average)
sd_quality_r6q2=sd(rank6_q2$average)
mean_quality_r6q2=mean(rank6_q2$average)
sd_quality_r6q3=sd(rank6_q3$average)
mean_quality_r6q3=mean(rank6_q3$average)
sd_quality_r6q4=sd(rank6_q4$average)
mean_quality_r6q4=mean(rank6_q4$average)

sd_distance_r6q1=sd(rank6_q1$distance)
mean_distance_r6q1=mean(rank6_q1$distance)
sd_distance_r6q2=sd(rank6_q2$distance)
mean_distance_r6q2=mean(rank6_q2$distance)
sd_distance_r6q3=sd(rank6_q3$distance)
mean_distance_r6q3=mean(rank6_q3$distance)
sd_distance_r6q4=sd(rank6_q4$distance)
mean_distance_r6q4=mean(rank6_q4$distance)

mean_sd_table_quantile_1=cbind(c(sd_cutoff_r1q1, sd_cutoff_r2q1, sd_cutoff_r3q1, sd_cutoff_r4q1, sd_cutoff_r5q1, sd_cutoff_r6q1), c(mean_cutoff_r1q1, mean_cutoff_r2q1, mean_cutoff_r3q1, mean_cutoff_r4q1, mean_cutoff_r5q1, mean_cutoff_r6q1), c(sd_quality_r1q1, sd_quality_r2q1, sd_quality_r3q1, sd_quality_r4q1, sd_quality_r5q1, sd_quality_r6q1), c(mean_quality_r1q1, mean_quality_r2q1, mean_quality_r3q1, mean_quality_r4q1, mean_quality_r5q1, mean_quality_r6q1), c(sd_distance_r1q1, sd_distance_r2q1, sd_distance_r3q1, sd_distance_r4q1, sd_distance_r5q1, sd_distance_r6q1), c(mean_distance_r1q1, mean_distance_r2q1, mean_distance_r3q1, mean_distance_r4q1, mean_distance_r5q1, mean_distance_r6q1))

colnames(mean_sd_table_quantile_1) = c("sd_cutoff", "mean_cutoff","sd_quality","mean_quality", "sd_distance", "mean_distance")
rownames(mean_sd_table_quantile_1) = c("choice1", "choice2","choice3","choice4", "choice5", "choice6")

mean_sd_table_quantile_2=cbind(c(sd_cutoff_r1q2, sd_cutoff_r2q2, sd_cutoff_r3q2, sd_cutoff_r4q2, sd_cutoff_r5q2, sd_cutoff_r6q2), c(mean_cutoff_r1q2, mean_cutoff_r2q2, mean_cutoff_r3q2, mean_cutoff_r4q2, mean_cutoff_r5q2, mean_cutoff_r6q2), c(sd_quality_r1q2, sd_quality_r2q2, sd_quality_r3q2, sd_quality_r4q2, sd_quality_r5q2, sd_quality_r6q2), c(mean_quality_r1q2, mean_quality_r2q2, mean_quality_r3q2, mean_quality_r4q2, mean_quality_r5q2, mean_quality_r6q2), c(sd_distance_r1q2, sd_distance_r2q2, sd_distance_r3q2, sd_distance_r4q2, sd_distance_r5q2, sd_distance_r6q2), c(mean_distance_r1q2, mean_distance_r2q2, mean_distance_r3q2, mean_distance_r4q2, mean_distance_r5q2, mean_distance_r6q2))

colnames(mean_sd_table_quantile_2) = c("sd_cutoff", "mean_cutoff","sd_quality","mean_quality", "sd_distance", "mean_distance")
rownames(mean_sd_table_quantile_2) = c("choice1", "choice2","choice3","choice4", "choice5", "choice6")

mean_sd_table_quantile_3=cbind(c(sd_cutoff_r1q3, sd_cutoff_r2q3, sd_cutoff_r3q3, sd_cutoff_r4q3, sd_cutoff_r5q3, sd_cutoff_r6q3), c(mean_cutoff_r1q3, mean_cutoff_r2q3, mean_cutoff_r3q3, mean_cutoff_r4q3, mean_cutoff_r5q3, mean_cutoff_r6q3), c(sd_quality_r1q3, sd_quality_r2q3, sd_quality_r3q3, sd_quality_r4q3, sd_quality_r5q3, sd_quality_r6q3), c(mean_quality_r1q3, mean_quality_r2q3, mean_quality_r3q3, mean_quality_r4q3, mean_quality_r5q3, mean_quality_r6q3), c(sd_distance_r1q3, sd_distance_r2q3, sd_distance_r3q3, sd_distance_r4q3, sd_distance_r5q3, sd_distance_r6q3), c(mean_distance_r1q3, mean_distance_r2q3, mean_distance_r3q3, mean_distance_r4q3, mean_distance_r5q3, mean_distance_r6q3))

colnames(mean_sd_table_quantile_3) = c("sd_cutoff", "mean_cutoff","sd_quality","mean_quality", "sd_distance", "mean_distance")
rownames(mean_sd_table_quantile_3) = c("choice1", "choice2","choice3","choice4", "choice5", "choice6")

mean_sd_table_quantile_4=cbind(c(sd_cutoff_r1q4, sd_cutoff_r2q4, sd_cutoff_r3q4, sd_cutoff_r4q4, sd_cutoff_r5q4, sd_cutoff_r6q4), c(mean_cutoff_r1q4, mean_cutoff_r2q4, mean_cutoff_r3q4, mean_cutoff_r4q4, mean_cutoff_r5q4, mean_cutoff_r6q4), c(sd_quality_r1q4, sd_quality_r2q4, sd_quality_r3q4, sd_quality_r4q4, sd_quality_r5q4, sd_quality_r6q4), c(mean_quality_r1q4, mean_quality_r2q4, mean_quality_r3q4, mean_quality_r4q4, mean_quality_r5q4, mean_quality_r6q4), c(sd_distance_r1q4, sd_distance_r2q4, sd_distance_r3q4, sd_distance_r4q4, sd_distance_r5q4, sd_distance_r6q4), c(mean_distance_r1q4, mean_distance_r2q4, mean_distance_r3q4, mean_distance_r4q4, mean_distance_r5q4, mean_distance_r6q4))

colnames(mean_sd_table_quantile_4) = c("sd_cutoff", "mean_cutoff","sd_quality","mean_quality", "sd_distance", "mean_distance")
rownames(mean_sd_table_quantile_4) = c("choice1", "choice2","choice3","choice4", "choice5", "choice6")

mean_sd_table
mean_sd_table_quantile_1
mean_sd_table_quantile_2
mean_sd_table_quantile_3
mean_sd_table_quantile_4



```








# Exercise 5
```{r echo=TRUE, results="show"}
#Exercise 5 Data Creation----
set.seed(100)
x1=runif(10000, 1, 3)
x2=rgamma(10000, 3, scale=2)
x3=rbinom(10000, 1, prob=0.3)
e=rnorm(10000, 2, 1)

y=0.5+1.2*x1-0.9*x2+0.1*x3+e
#hist(y)
ydum=as.numeric((y>mean(y)))
```
# Exercise 6
```{r echo=TRUE, results="show"}
#Exercise 6 OLS----
cor(y, x1)

X=cbind(1, x1, x2, x3)
b=solve(t(X)%*%X)%*%(t(X)%*%y)
y_hat=X%*%b
e_hat_2=(y-X%*%b)^2
b
Var=sum(e_hat_2)/(nrow(X)-ncol(X))*solve(t(X)%*%X)

standard_error=sqrt(diag(Var))
standard_error
#check
#model=lm(y~X)
#summary(model)
```
| Correlation of y and X_1 is shown above.
| Coefficient of y on x_1 is 1.24059, it's not very different from 1.2, and is statistically significant.
| Coefficients of y on X is shown above as b.
| Standard Error is shown above
# Exercise 7
```{r echo=TRUE, results="show"}
#Exercise 7 Discrete Choice----

#probit
reg1 = glm(ydum~x1+x2+x3,family = binomial(link = "probit"))
summary(reg1)

flike = function(par,x1,x2,x3,ydum)
{
  xbeta           = par[1] + par[2]*x1 + par[3]*x2 + par[4]*x3
  pr              = pnorm(xbeta)
  #  pr              = exp(beta)/(1+exp(beta)) logit
  pr[pr>0.999999] = 0.999999
  pr[pr<0.000001] = 0.000001
  like           = ydum*log(pr) + (1-ydum)*log(1-pr)
  return(-sum(like))
}

start = runif(4)
res  = optim(start,fn=flike,method="BFGS",control=list(trace=6,REPORT=1,maxit=1000),x1=x1,x2=x2,x3=x3,ydum=ydum,hessian=TRUE)

fisher_info_probit = solve(res$hessian)       # standard formula is -res$hessian but flike is return -like
prop_sigma_probit  = sqrt(diag(fisher_info_probit))
#prop_sigma


#logit
reg2 = glm(ydum~x1+x2+x3,family = binomial(link = "logit"))
summary(reg2)

flike_logit = function(par,x1,x2,x3,ydum)
{
  xbeta_logit           = par[1] + par[2]*x1 + par[3]*x2 + par[4]*x3
  pr_logit              = exp(xbeta_logit)/(1+exp(xbeta_logit))
  pr_logit[pr_logit>0.999999] = 0.999999
  pr_logit[pr_logit<0.000001] = 0.000001
  like_logit           = ydum*log(pr_logit) + (1-ydum)*log(1-pr_logit)
  return(-sum(like_logit))
}

start_logit = runif(4)
res_logit  = optim(start_logit,fn=flike_logit,method="BFGS",control=list(trace=6,REPORT=1,maxit=1000),x1=x1,x2=x2,x3=x3,ydum=ydum,hessian=TRUE)

fisher_info_logit = solve(res_logit$hessian)       # standard formula is -res$hessian but flike is return -like
prop_sigma_logit  = sqrt(diag(fisher_info_logit))
#prop_sigma_logit

#Linear
linear=lm(ydum~x1+x2+x3)
summary(linear)




est_probit = cbind(summary(reg1)$coefficients[, 1],summary(reg1)$coefficients[, 2],res$par,prop_sigma_probit)
colnames(est_probit) = c("Probit : est","Probit :se","Probit: own : est","Probit: own :se")
est_probit

est_logit = cbind(summary(reg2)$coefficients[, 1],summary(reg2)$coefficients[, 2],res_logit$par,prop_sigma_logit)
colnames(est_logit) = c("Logit : est","Logit :se","Logit: own : est","Logit: own :se")
est_logit
```
| Table est_probit shows the point estimation and SE of porbit model and optimization results.
| Table est_logit shows the point estimation and SE of logit model and optimization results.
| The value of the point estimation of probit and logit model is quite different, which doesn't matter because the point estimation doesn't tell us the marginal effect.
| The sign of the point estimation of probit and logit model is the same, which makes sense, because although we cannot interpret the magnitude, we can interpret if it's more likely or less likely for the dependent variable to be 1 in this case keeping all else constant.
| In terms of significance, by calculating the p value, point estimation of probit and logit model is significant. In linear model, however, coefficient for x3 is not significant, coefficients for x1&x2 are significant though.
# Exercise 8
```{r echo=TRUE, results="show"}
#Exercise 8 Marginal Effects----
#Probit Model Average Marginal Effects
probit_scalar=mean(dnorm(X%*%res$par))
probit_margin=as.matrix(probit_scalar*res$par)
probit_margin
#Logit Model Average Marginal Effects
logit_scalar=mean(dlogis(X%*%res_logit$par))
logit_margin=as.matrix(logit_scalar*res_logit$par)
logit_margin
#SE
X_all=as.data.frame(cbind(ydum,x1,x2,x3))
x_mean=as.matrix(colMeans(X_all))
mat=as.matrix(res$par)
lll=length(res$par)
xb=t(x_mean)%*%mat
vcv=solve(res$hessian)
gr=apply(cbind(1, x1, x2, x3), 1, function(x){
as.numeric(as.numeric(dnorm(x %*% mat))*(diag(lll) - as.numeric(x %*% mat)*(mat %*% t(x))))
})
gr = matrix(apply(gr,1,mean),nrow=lll)
Probit_marg_SE = sqrt(diag(gr %*% vcv %*% t(gr)))
Probit_marg_SE

X_all=as.data.frame(cbind(ydum,x1,x2,x3))
x_mean=as.matrix(colMeans(X_all))
mat=as.matrix(res_logit$par)
lll=length(res_logit$par)
xb=t(x_mean)%*%mat
vcv=solve(res_logit$hessian)
gr = apply(cbind(1,x1,x2,x3), 1, function(x){
as.numeric(as.numeric(plogis(x %*% mat)*(1-plogis(x %*% mat)))*
(diag(lll) - (1 - 2*as.numeric(plogis(x %*% mat)))*(mat %*% t(x))))
})
gr = matrix(apply(gr,1,mean),nrow=lll)
Logit_marg_SE = sqrt(diag(gr %*% vcv %*% t(gr)))
Logit_marg_SE

```
| Marginal effect of probit and logit model is calculated in probit_margin and logit_margin.
| Standard error of probit and logit marginal effects is shown in Probit_marg_SE and Logit_marg_SE.













