---
title: "R Notebook"
output: html_notebook
---




```{r, include}
setwd("D:/wtk/work_qq/R/y2104401")
require(tidyverse)
require(ggplot2)
require(dplyr)
require(plm)
require(fixest)
```



```{r}
population <- read.csv("population.csv")
crime_long <- read.csv("crime_long.csv")
officers <- read.csv("officers.csv")

crime_long <- crime_long %>% 
  group_by(crime_month, district, crime_type) %>%
  summarise(crimes=sum(crimes)) %>% tibble()
```


# exercise 2

## Calculate total crime per month and plot the time series of crime.
```{r}
crime_long$crime_month <- as.Date(crime_long$crime_month)
```


```{r}
crime_long %>%
  select(crime_month, crimes) %>%
  group_by(by=crime_month) %>% summarise(crimes.sum=sum(crimes)) %>%
  mutate(month=by) %>%
  ggplot(aes(month, crimes.sum)) +
  geom_line() +
  xlab("time") +
  ylab("crimes")
```


## Merge the two datasets by districts-units and period.

```{r}
population$month <- as.Date(population$month)
```



```{r}
population.crime <- population %>% left_join(crime_long, by=c("month"="crime_month", 
                                          "district"="district"))

population.crime$white.share = round(population.crime$tot_white/population.crime$tot_pop, 3)
population.crime$hisp.share = round(population.crime$tot_hisp / population.crime$tot_pop, 3)
population.crime$black.share = round(population.crime$tot_black / population.crime$tot_pop, 3)


population.crime <- population.crime %>% drop_na() %>%
  select(-tot_pop, -tot_white, -tot_black, -tot_hisp) %>%
  spread(crime_type, crimes) %>% 
  mutate(tot.crimes=drug + other + property + violent) %>%
  select(-c("other", "drug", "period")) %>%
  select(district, everything()) %>% arrange(district, month)

```


```{r}
head(population.crime)
```



# Exercise 3 Panel Data: Introduction

```{r}
officers <- read.csv("officers.csv")
officers$month <- as.Date(officers$month, format = "%m/%d/%Y")

officers.crime <- officers %>% 
  left_join(population.crime, by=c("month"="month", "unit"="district")) %>%
  drop_na()
head(officers.crime)
```






```{r}
# ols regression
officers.crime.3.lm <- lm(arrest~tenure+p50_inc+white.share+hisp.share+black.share+tot.crimes+property+violent - 1, officers.crime) #
## coefficients of ols
### tenture
officers.crime.3.lm$coefficients[1]
### others
officers.crime.3.lm$coefficients[2:length(officers.crime.3.lm$coefficients)]

# panel regression, pooling, used to verify the result above.
officers.crime.3.plm <- plm(arrest~tenure+p50_inc+white.share+hisp.share+black.share+tot.crimes+property+violent - 1, officers.crime,model = "pooling", index = c("NUID", "month")) #

## coefficients of plm, pooling

### tenture
officers.crime.3.plm$coefficients[1]
### others
officers.crime.3.plm$coefficients[2:length(officers.crime.3.lm$coefficients)]
```

# Exercise 4 Panel Data: More controls




```{r}
# ols
officers.crime.4.lm <- lm(arrest~tenure+p50_inc+white.share+hisp.share+black.share+tot.crimes+property+violent +factor(month) + factor(unit) - 1, officers.crime) #

oc.4.lm.f <- grep("factor", as.vector(names(officers.crime.4.lm$coefficients)))
## coefficients of ols
### tenure
officers.crime.4.lm$coefficients[1]
### Z
officers.crime.4.lm$coefficients[2:oc.4.lm.f[1]-1]
### district fixed effects
district.f <- grep("factor.unit.", as.vector(names(officers.crime.4.lm$coefficients)))
officers.crime.4.lm$coefficients[district.f]
### year and month fixed effects
time.f <- grep("factor.month.", as.vector(names(officers.crime.4.lm$coefficients)))
officers.crime.4.lm$coefficients[time.f]
```


# Exercise 5 Panel Data: Individual xed eects


```{r}
# within
officers.crime.5.plm <- plm(arrest~tenure+p50_inc+white.share+hisp.share+black.share+tot.crimes+property+violent+factor(unit), 
                              data= officers.crime, model = "within", index =  c("NUID", "month"))
```

```{r}
# between
officers.crime.5.plm.b <- plm(arrest~tenure+p50_inc+white.share+hisp.share+black.share+tot.crimes+property+violent+factor(unit),
                              data= officers.crime, model = "between", index =  c("NUID", "month"))
```



```{r}
# first difference
officers.crime.5.plm.d <- plm(arrest~tenure+p50_inc+white.share+hisp.share+black.share+tot.crimes+property+violent+factor(unit),
                              data= officers.crime, model = "fd", index =  c("NUID", "month"))
```


```{r}
coef.within.tenure <- as.vector(officers.crime.5.plm$coefficients[1])
coef.between.tenure <- as.vector(officers.crime.5.plm.b$coefficients[2])
coef.fd.tenure <- as.vector(officers.crime.5.plm.d$coefficients[2])
pander::pandoc.table(data.frame(coef.within.tenure ,coef.between.tenure ,coef.fd.tenure))
```
As we known, $\beta_{within, tenture}$ is `r coef.within.tenure`,  $\beta_{between, tenture}$ is `r coef.between.tenure`,
 $\beta_{fd, tenture}$ is `r coef.fd.tenure`. so, estimated effect by between and fd both are positive effects on dependent variable, estimated effect by within is negative. And estimated level by first difference is the largest.

## b


```{r}
officers.crime.5.p <- feols(arrest~tenure+p50_inc+white.share+hisp.share+black.share+tot.crimes | month + unit + NUID, officers.crime, panel.id = c("NUID", "month"))
etable(officers.crime.5.p)
```


```{r}
officers.crime.5.p.coef <- fixef(officers.crime.5.p)
summary(officers.crime.5.p.coef)
```



one-step GMM:

$$
\hat\beta_{2SLS}=[X'Z(Z'Z)^{-1}Z'X]^{-1}X'Z(Z'Z)^{-1}Z'y
$$
if X = Z, it is just estimator of OLS.

